Novel Sampling Using Stochastic Gradient Descent

### 2016-12-05 poster/main.tex
update the mid column

\newcommand{\bx}{\mathbf{x}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\D}{\mathscr{D}}
\newcommand{\N}{\mathscr{N}}
\newcommand{\by}{\mathbf{y}}
\newcommand{\var}{\mathrm{var}}
\DeclareMathOperator*{\argmin}{\arg\!\min}

@@

\begin{block}{SGD Sampling Algorithm}
Suppose we would like to sample from a distribution $p(x) \propto f(x)$. 
\begin{algorithm}[H]
{\small
Given $f(x)$, 
\begin{algorithmic}\caption{Stochastic gradient descent (oracle)}\label{Alg:SGD}
\STATE {\bf Initialize} $ \bx^0$ arbitrarily;
\WHILE{burn-in period}
\STATE draw $a^t$ from Gaussian distribution $\N(\nabla \log f(\bx^k), \Sigma^{t})$.
\STATE
$\bx^{k+1} \gets \bx^k + (2\Sigma^t)^{-1} a^t$;
\STATE $k\gets k+1$;
\ENDWHILE
\STATE return $\bx^k$.
\end{algorithmic}}
\end{algorithm}

@@

%----------------------------------------------------------------------------------------
%	MATERIALS
%----------------------------------------------------------------------------------------

\begin{block}{Stationary Distribution}
\begin{conjecture}
For any $p(\bx)\in \D(\R^p)$, it is the stationary distribution of iterates in Algorithm 1 with parameter $(-\log p(\bx), \epsilon, 2\epsilon)$.
\end{conjecture}

@@

\begin{block}{SGD As Posterior Sampler}

\begin{theorem}[Central limit theorem]
Let $K = \alpha^{-1} \var \log P(X_1|\theta^t)$,
\[
\sum_{i=1}^K (K^{-1}\nabla \log P(X_{\eta_i}|\theta^t) - \log P(\theta^t)) - (n^{-1} \sum_{i=1}^n \nabla \log P(X_i|\theta^t) - \log P(\theta^t)) \sim \N(0, \alpha).
\]
\end{theorem}

A toy example showing SGD works.

@@

\begin{block}{Upper Bound On Mixing Time}

Maecenas ultricies feugiat velit non mattis. Fusce tempus arcu id ligula varius dictum.
\begin{itemize}
\item Curabitur pellentesque dignissim
\item Eu facilisis est tempus quis
\item Duis porta consequat lorem
\end{itemize}

>>>>

\newcommand{\bx}{\mathbf{x}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\D}{\mathscr{D}}
\newcommand{\N}{\mathscr{N}}
\newcommand{\by}{\mathbf{y}}
\newcommand{\var}{\mathrm{var}}
\newcommand{\cov}{\mathrm{cov}}
\DeclareMathOperator*{\argmin}{\arg\!\min}

@@

\begin{block}{SGD Sampling Algorithm}
Suppose we would like to sample from a distribution $p(\bx)\in \D(\R^p) \propto f(\bx)$. 
\begin{algorithm}[H]
{\small
Given $f(x)$, 
\begin{algorithmic}\caption{Stochastic gradient descent (oracle)}\label{Alg:SGD}
\STATE {\bf Initialize} $ \bx^0$ arbitrarily;
\WHILE{burn-in period}
\STATE draw $a^t$ from Gaussian distribution $\N(\nabla \log f(\bx^k), \Sigma^{t})$.
\STATE
$\bx^{k+1} \gets \bx^k + (2\Sigma^t)^{-1} a^t$;
\STATE $k\gets k+1$;
\ENDWHILE
\STATE return $\bx^k$.
\end{algorithmic}}
\end{algorithm}

@@

%----------------------------------------------------------------------------------------
%	MATERIALS
%----------------------------------------------------------------------------------------
\begin{block}{Stationary Distribution}
Will Algorithm 1 converge to the desired stationary distribution? Denote $p(\bx^{t+1}|\bx^t)$ to be the density function to obtain $\bx^{t+1}$ from $\bx^{t}$. 
The detailed balance condition holds approximately when $\|(\Sigma^t)^{-1}\|$ is small, i.e.,
\[
f(\bx^{t})p(\bx^{t+1}|\bx^t) = f(\bx^{t+1})p(\bx^{t}|\bx^{t+1}) ( 1 + O(\|(\Sigma^t)^{-1}\|^2)).
\]
That implies $f(\cdot)$ is indeed the stationary distribution of Algorithm 1. Simulation results can be found in Figure 1.
\end{block}

@@

\begin{block}{SGD As Posterior Sampler}
When $p(\theta)$ is the posterior distribution of some parameters given a set of data $X_1,\ldots, X_n$, i.e.,
\[
p(\theta) = \prod_{i=1}^n P(X_i|\theta) P(\theta),
\]
The noisy gradient step in Algorithm 1 can be estimated by subsampling:
\begin{theorem}[Central limit theorem]
Let $K$ be the subsampling size and $\eta_1,\ldots, \eta_K$ be samples drawing from $1,\ldots, n$ with replacement, then
\[
K^{-1}\sum_{i=1}^K (\nabla \log P(X_{\eta_i}|\theta^t) - \log P(\theta^t)) \sim \N(n^{-1} \nabla \log p(\theta) , K^{-1}\cov \nabla \log p(\theta)).
\]
\end{theorem}

See Figure 2 for the simulation result.

@@

\begin{block}{Upper Bound On Mixing Time}
Mixing time correspon
